Consensus cell types across samples from brain and CNS tumors

### 2025-02-12 analyses/cell-type-consensus/exploratory-notebooks/04-brain-and-cns-consensus-celltypes.Rmd
adjust HGG and myeloid groupings

- Low-grade glioma: Any samples labeled with "Low-grade glioma" or are part of `SCPCP000002` which are categorized as LGG tumors. 
- High-grade glioma: Any samples labeled with "High-grade glioma" or are part of `SCPCP000001` which are categorized as HGG tumors. 
- Glioblastoma
- Pilocytic astrocytoma
- Ganglioglioma 
- ATRT 
- Other: All samples that do not fall into the above categories.

@@

```{r}
# assign diagnosis groups to use for faceting plots 
all_results_df <- all_results_df |> 
  dplyr::mutate(
    grouped_diagnosis = dplyr::case_when(
      diagnosis %in% c("Low-grade glioma", "Glioblastoma", "Pilocytic astrocytoma", "Ganglioglioma", "High-grade glioma", "Atypical teratoid rhabdoid tumor", "Non-cancerous") ~ diagnosis,
      project_id == "SCPCP000001" ~ "High-grade glioma",
      project_id == "SCPCP000002" ~ "Low-grade glioma",
      .default =  "other"
    ) |>
      forcats::fct_relevel(c("other", "Non-cancerous"), after = Inf)
  )
```

@@

```{r}
# t cell subtypes
t_cells <- unique_celltypes |> 
  stringr::str_subset("T")

# macro/mono lineage 
macro_mono_cells <- c(
  "macrophage", 
  "monocyte",
  "mononuclear phagocyte", 
  "dendritic cell"
)

# create broad category for T cell vs. macro/mono 
immune_results_df <- immune_results_df |> 
  dplyr::mutate(
    immune_broad_category = dplyr::case_when(
      consensus_annotation %in% t_cells ~ "T cell", 
      consensus_annotation %in% macro_mono_cells ~ "Macro/mono",
      consensus_annotation == "Unknown" ~ "Unknown",
      .default = "All remaining cell types"
    )
  )

# define colors 
immune_category_colors <- c(
  "T cell" = all_celltype_colors[["T cell"]],
  "Macro/mono" = all_celltype_colors[["macrophage"]],
  "All remaining cell types" = "grey60",
  "Unknown" = "grey90"
)
```

>>>>

- Low-grade glioma: Any samples labeled with "Low-grade glioma" or are part of `SCPCP000002` which are categorized as LGG tumors. 
- High-grade glioma: Any samples labeled with "High-grade glioma", "Diffuse midline glioma", or are part of `SCPCP000001` which are categorized as HGG tumors. 
- Glioblastoma
- Pilocytic astrocytoma
- Ganglioglioma 
- ATRT 
- Other: All samples that do not fall into the above categories.

@@

```{r}
# assign diagnosis groups to use for faceting plots 
all_results_df <- all_results_df |> 
  dplyr::mutate(
    grouped_diagnosis = dplyr::case_when(
      diagnosis %in% c("Low-grade glioma", "Glioblastoma", "Pilocytic astrocytoma", "Ganglioglioma", "High-grade glioma", "Atypical teratoid rhabdoid tumor", "Non-cancerous") ~ diagnosis,
      project_id == "SCPCP000001" | diagnosis == "Diffuse midline glioma" ~ "High-grade glioma",
      project_id == "SCPCP000002" ~ "Low-grade glioma",
      .default =  "other"
    ) |>
      forcats::fct_relevel(c("other", "Non-cancerous"), after = Inf)
  )
```

@@

```{r}
# t cell subtypes
t_cells <- unique_celltypes |> 
  stringr::str_subset("T cell")

# macro/mono lineage 
myeloid_cells <- c(
  "macrophage", 
  "monocyte",
  "mononuclear phagocyte", 
  "myeloid leukocyte",
  "neutrophil"
)

# create broad category for T cell vs. macro/mono 
immune_results_df <- immune_results_df |> 
  dplyr::mutate(
    immune_broad_category = dplyr::case_when(
      consensus_annotation %in% t_cells ~ "T cell", 
      consensus_annotation %in% myeloid_cells ~ "Myeloid",
      consensus_annotation == "Unknown" ~ "Unknown",
      .default = "All remaining cell types"
    )
  )

# define colors 
immune_category_colors <- c(
  "T cell" = all_celltype_colors[["T cell"]],
  "Myeloid" = all_celltype_colors[["macrophage"]],
  "All remaining cell types" = "grey60",
  "Unknown" = "grey90"
)
```
